Read Me

### 2023-04-30 code/analysis_code/3_model_alt.Rmd
fixed pca and models

```{r}
df <- readRDS(here("data/processed_data/df_scaled_alt.rds"))
us_nabs <- readRDS(here("data/processed_data/us_nabs.rds"))

@@

Setting up functions for models
```{r}
 #multiple linear regression
mult_function <- function(df) {
  mult_race <- glm(MRR ~ 
                        `Percent Living in Poverty` +
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old` +
                        `Median Household Income`  +
                     #   `Percent Living with Severe Housing Cost Burden` +
                        `Percent Living with Food Insecurity`, 
                     data = df,
                     family = "gaussian")
  print(summary(mult_race))
    ##pull fitted values
df$mult_fit <- mult_race$fitted.values
return(df)
}

## function to fit GLM and pull AIC
mult_aic <- function(df) {
  mult_race <- glm(MRR ~ 
                        `Percent Living in Poverty` +
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old` +
                        `Median Household Income`  +
                       # `Percent Living with Severe Housing Cost Burden` +
                        `Percent Living with Food Insecurity`,
                     data = df,
                     family = "gaussian")
aic <- AIC(mult_race)

@@

pois_race <- glm(`Death Cases` ~ Race +
                        `Percent Living in Poverty` +
                       # `Gini Coefficient` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old` +
                        `Median Household Income`  +
                        `Percent Living with Food Insecurity` +
                        #`Percent Living with Severe Housing Cost Burden` +
                     offset(log(`Population`)),
                     data = df,
                     family = poisson(link = "log"))
  print(summary(pois_race))

@@

## function to fit GLM and pull AIC
pois_aic <- function(df) {
  pois_race <- glm(`Death Cases` ~  Race +
                        `Percent Living in Poverty` +
                      #  `Gini Coefficient` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old` +
                        `Median Household Income`  +
                        `Percent Living with Food Insecurity` +
                       # `Percent Living with Severe Housing Cost Burden` +
                     offset(log(`Population`)),
                     data = df,
                     family = poisson(link = "log"))
aic <- AIC(pois_race)
return(aic)
}
```

@@

```{r}
  #Negative binomial
nb_function  <- function(df) {
  neg_bin_race <- glm.nb(`Death Cases` ~ Race +
                        `Percent Living in Poverty` +
                       # `Gini Coefficient` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old` +
                        `Median Household Income`  +
                        `Percent Living with Food Insecurity`  +
                       # `Percent Living with Severe Housing Cost Burden` +
                     offset(log(`Population`)),
                            data = df)
  print(summary(neg_bin_race))

@@

## function to fit GLM and pull AIC
nb_aic <- function(df) {
  neg_bin_race <- glm.nb(`Death Cases` ~ Race +
                        `Percent Living in Poverty` +
                       # `Gini Coefficient` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old` +
                        `Median Household Income`  +
                        `Percent Living with Food Insecurity`  +
                        #`Percent Living with Severe Housing Cost Burden` +
                     offset(log(`Population`)),
                            data = df)
aic <- AIC(neg_bin_race)

@@

  car_race <- spautolm(MRR ~                      
                        `Percent Living in Poverty` +
                       `Median Household Income` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old` +
                     #   `Percent Unemployed`  +
                     #   `Percent Living with Severe Housing Cost Burden` +
                        `Percent Living with Food Insecurity`, 
                           data = df,
                           listw = nablist_race_df,
                           family = "CAR")
  print(summary(car_race))

@@

car_aic <- function(df) {
  h_race_df <- matrix(0, nrow(df), nrow(df))
  h_race_df[cbind(match(us_nabs$StateCode, df$`State Abbreviation`),
                  match(us_nabs$NeighborStateCode, df$`State Abbreviation`))] <- 1

@@

  car_race <- spautolm(MRR ~                      
                        `Percent Living in Poverty` +
                       `Median Household Income` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old` +
                   #     `Percent Unemployed`  +
                      #  `Percent Living with Severe Housing Cost Burden` +
                        `Percent Living with Food Insecurity`, 
                           data = df,
                           listw = nablist_race_df,
                           family = "CAR")
aic <- AIC(car_race)
return(aic)

@@

Organizing mod_AIC into new dataframe
```{r}
aic_df <- data.frame(AIC = c(df_mult, df_pois, df_nb, df_car),
                     Model = c("GLM", "Poisson", "Negative Binomial", "CAR"),
                     Outcome = c("Ratio", "Count", "Count", "Ratio"))
aic_df <- aic_df %>% mutate(AIC  = round(AIC, 2))
```

```{r}
saveRDS(df, here("data/processed_data/df_fit_alt.rds"))

saveRDS(aic_df, here("data/processed_data/aic_alt.rds"))
```

>>>>

```{r}
df <- readRDS(here("data/processed_data/race_alt.rds"))
prin_comp1 <- readRDS(here("data/processed_data/prin_comp1.rds"))
us_nabs <- readRDS(here("data/processed_data/us_nabs.rds"))

@@

Setting up functions for models
```{r}
 #multiple linear regression
mult_function <- function(df) {
  mult_race <- glm(MRR ~ 
                        prin_comp1,
                     data = df,
                     family = "gaussian")
  print(summary(mult_race))
    ##pull fitted values
df$mult_fit <- mult_race$fitted.values
return(df)
}

## function to fit GLM and pull AIC
mult_aic <- function(df) {
  mult_race <- glm(MRR ~ 
                        prin_comp1,
                     data = df,
                     family = "gaussian")
aic <- AIC(mult_race)

@@

pois_race <- glm(`Death Cases` ~ Race + region +
                        prin_comp1 +
                     offset(log(`Population`)),
                     data = df,
                     family = poisson(link = "log"))
  print(summary(pois_race))

@@

## function to fit Poisson and pull AIC
pois_aic <- function(df) {
  pois_race <- glm(`Death Cases` ~  Race + region +
                         prin_comp1  +
                     offset(log(`Population`)),
                     data = df,
                     family = poisson(link = "log"))
aic <- AIC(pois_race)
return(aic)
}
```

@@

```{r}
  #Negative binomial
nb_function  <- function(df) {
  neg_bin_race <- glm.nb(`Death Cases` ~  Race + region +
                         prin_comp1  +
                     offset(log(`Population`)),
                            data = df)
  print(summary(neg_bin_race))

@@

## function to fit nb and pull AIC
nb_aic <- function(df) {
  neg_bin_race <- glm.nb(`Death Cases` ~ Race + region +
                         prin_comp1  +
                     offset(log(`Population`)),
                            data = df)
aic <- AIC(neg_bin_race)

@@

  car_race <- spautolm(MRR ~   prin_comp1, 
                           data = df,
                           listw = nablist_race_df,
                           family = "CAR")
  print(summary(car_race))

@@

##pulling AIC from CAR model
car_aic <- function(df) {
  h_race_df <- matrix(0, nrow(df), nrow(df))
  h_race_df[cbind(match(us_nabs$StateCode, df$`State Abbreviation`),
                  match(us_nabs$NeighborStateCode, df$`State Abbreviation`))] <- 1

@@

  car_race <- spautolm(MRR ~                      
                        prin_comp1,
                           data = df,
                           listw = nablist_race_df,
                           family = "CAR")
aic <- AIC(car_race)
return(aic)

@@

Organizing mod_AIC into new dataframe
```{r}
aic_df <- data.frame(AIC = c(df_mult, df_pois, df_nb, df_car),
                     Model = c("Multiple Linear Regression", "Poisson", "Negative Binomial", "Conditional Autoregressive"),
                     Outcome = c("Ratio", "Count", "Count", "Ratio"))

aic_df <- aic_df %>%
  mutate(AIC  = round(AIC, 2))

```{r}
##Making kable table of model results
library(kableExtra)
library(knitr)
table_fit <- aic_df %>%
  select(c(Model, Outcome, AIC)) %>% 
  kable(caption = "<center>Model Fit</center>",
        align = "c") %>%
  kable_minimal() %>%
  kable_styling(font_size = 30) %>%
gsub("font-size: initial !important;",
"font-size: 30pt !important;",
.)  %>% 
  landscape()

table_fit %>%  
  save_kable(here("results/table_fit.png"))

```{r}
saveRDS(df, here("data/processed_data/df_fit_alt.rds"))
```

Testing for spatial association. 
```{r}
#setting up matrix
  h_race_df <- matrix(0, nrow(df), nrow(df))
  h_race_df[cbind(match(us_nabs$StateCode, df$`State Abbreviation`),
  match(us_nabs$NeighborStateCode, df$`State Abbreviation`))] <- 1

#creating neighbor list 
  nablist_race_df <- mat2listw(h_race_df, style = "W")

# Using testing for spatial association using Moran's test (global)
  print(moran.test(df$MRR, nablist_race_df, zero.policy = TRUE, randomisation = FALSE))
# plot Moran's I
  moran.plot(df$MRR, nablist_race_df, zero.policy = TRUE)
  #save plot
png(here("results/moran_mrr.png"))
  moran.plot(df$MRR, nablist_race_df, zero.policy = TRUE)
dev.off()

  # calculate and print LISA statistics (local)
lisa_race <- localmoran(df$MRR, nablist_race_df, zero.policy=TRUE)
  summary(lisa_race[,1]) 
```
